A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
stuff

\lstset{ 
 language=R,              % the language of the code
 basicstyle=\small\ttfamily,    % the size of the fonts
% numbers=left,            % where to put the line-numbers
 numberstyle=\color{Blue},% the style that is used for the line-numbers
 stepnumber=1,            % the step between two line-numbers. If it is 1,
                          % each line will be numbered
 numbersep=2pt,           % how far the line-numbers are from the code
 backgroundcolor=\color{white},  % choose the background color.
 showspaces=false,        % show spaces adding particular underscores
 showstringspaces=false,  % underline spaces within strings
 showtabs=false,          % show tabs within strings adding particular 
%frame=single,            % adds a frame around the code
 rulecolor=\color{black}, % if not set, the frame-color may be changed on
                          % line-breaks within not-black text (e.g. commens
                          % (green here))
 tabsize=2,               % sets default tabsize to 2 spaces
 captionpos=b,            % sets the caption-position to bottom
 breaklines=true,         % sets automatic line breaking
 breakatwhitespace=false, % sets if automatic breaks should only happen at 
 keywordstyle=\color{Blue},    % keyword style
 commentstyle=\color{YellowGreen},  % comment style
 stringstyle=\color{ForestGreen}    % string literal style
}

@@

\newcommand{\eval}{\texttt{eval}\xspace}
\newcommand{\parse}{\texttt{parse}\xspace}
\newcommand{\source}{\texttt{source}\xspace}
\newcommand{\local}{\texttt{local}\xspace}
\newcommand{\unlockBinding}{\texttt{unlockBinding}\xspace}
\newcommand{\substitute}{\texttt{substitute}\xspace}
\newcommand{\datatable}{\texttt{data.table}\xspace}
\newcommand{\mlogit}{\texttt{mlogit}\xspace}
\newcommand{\mboost}{\texttt{mboost}\xspace}
\newcommand{\metafor}{\texttt{metafor}\xspace}
\newcommand{\lavaan}{\texttt{lavaan}\xspace}
\newcommand{\mclust}{\texttt{mclust}\xspace}
\newcommand{\gamlss}{\texttt{gamlss}\xspace}
\newcommand{\ggproto}{\texttt{ggproto}\xspace}
\newcommand{\ggplot}{\texttt{ggplot2}\xspace}

\newcommand{\base}{\texttt{base}\xspace}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

\section{The Design of Eval}

@@

R supports reflection and meta-programming. The \c{substitute(e,envir)}
function yields the parse tree of the expression \c{e} after performing
substitutions defined by the bindings in \c{envir}.

\begin{lstlisting}
> substitute(expression(a + b), list(a = 1)))  
expression(1 + b)
\end{lstlisting}
\vspace{2mm}

\noindent
R allows programmatic manipulation of parse trees, which are themselves
first class objects. They are evaluated using the \c{eval(e,envir)}
function. Environment on the call stack can also be accessed.

There are four different kinds of \eval in the \emph{base} package:
\c{eval}, \c{evalq}, \c{eval.parent}, \c{local}.  The most important one is
\eval:

\begin{lstlisting}
  eval <- function(expr, envir=parent.frame(),
                   enclos=if (is.list(envir)||is.pairlist(envir)) parent.frame() else baseenv())  
	.Internal(eval(expr, envir, enclos))
\end{lstlisting}

The three other functions are defined using \eval.

\begin{lstlisting}
evalq <- function (expr, envir = parent.frame(), enclos = if (is.list(envir) || is.pairlist(envir)) parent.frame() else baseenv())  
	.Internal(eval(substitute(expr), envir, enclos))

eval.parent <- function (expr, n = 1) 
{
	p <- parent.frame(n + 1)
	eval(expr, p)
}

local <- function (expr, envir = new.env()) eval.parent(substitute(eval(quote(expr), envir)))
\end{lstlisting}

\c{expr} is the expression to be evaluated. It is not a string as in javascript. In order to evaluate a string, it should be first parsed using \c{parse(text = str)} to get an expression.

\eval makes it possible to specify the environment in which to evaluate the expression. \c{enclos} is an additional environment used for lookup when \c{envir} is a list, a pairlist or a dataframe. In that case, \c{envir} is copied into a temporary environment enclosed in \c{enclos}.

@@

Unlike CRAN, Kaggle is not a curated repository and therefore there are no
guarantees about the quality of the code. After downloading all of the
\CorpusKaggle kernels and extracting the R code from the various
formats\footnote{We use \texttt{rmarkdown} package to convert from Jupyer
  notebook to Rmarkdown as well as to convert from Rmarkdown to plain R}, we
found that \CorpusDuplicatedKaggle were whole-file duplicates. From
\CorpusRunnableKaggle kernels, \CorpusFailedKaggle failed to execute. Next to
various runtime exceptions, common problems were parse errors and misspelled
package names.

>>>>

\lstset{ 
 language=R,              % the language of the code
 basicstyle=\small\sffamily,    % the size of the fonts
% numbers=left,            % where to put the line-numbers
 numberstyle=\color{Blue},% the style that is used for the line-numbers
 stepnumber=1,            % the step between two line-numbers. If it is 1,
                          % each line will be numbered
 numbersep=2pt,           % how far the line-numbers are from the code
 backgroundcolor=\color{white},  % choose the background color.
 showspaces=false,        % show spaces adding particular underscores
 showstringspaces=false,  % underline spaces within strings
 showtabs=false,          % show tabs within strings adding particular 
%frame=single,            % adds a frame around the code
 rulecolor=\color{black}, % if not set, the frame-color may be changed on
                          % line-breaks within not-black text (e.g. commens
                          % (green here))
 tabsize=2,               % sets default tabsize to 2 spaces
 captionpos=b,            % sets the caption-position to bottom
 breaklines=true,         % sets automatic line breaking
 breakatwhitespace=false, % sets if automatic breaks should only happen at 
 keywordstyle=\color{Blue},    % keyword style
 commentstyle=\color{YellowGreen},  % comment style
 stringstyle=\color{ForestGreen}    % string literal style
}

@@

\newcommand{\eval}{\c{eval}\xspace}
\newcommand{\parse}{\c{parse}\xspace}
\newcommand{\source}{\c{source}\xspace}
\newcommand{\local}{\c{local}\xspace}
\newcommand{\unlockBinding}{\c{unlockBinding}\xspace}
\newcommand{\substitute}{\c{substitute}\xspace}
\newcommand{\datatable}{\c{data.table}\xspace}
\newcommand{\mlogit}{\c{mlogit}\xspace}
\newcommand{\mboost}{\c{mboost}\xspace}
\newcommand{\metafor}{\c{metafor}\xspace}
\newcommand{\lavaan}{\c{lavaan}\xspace}
\newcommand{\mclust}{\c{mclust}\xspace}
\newcommand{\gamlss}{\c{gamlss}\xspace}
\newcommand{\ggproto}{\c{ggproto}\xspace}
\newcommand{\ggplot}{\c{ggplot2}\xspace}

\newcommand{\base}{\c{base}\xspace}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

\section{R and The Design of Eval}

@@

R supports reflection and meta-programming. The function
\c{substitute(e,envir)} yields the parse tree of the expression \c{e} after
performing substitutions defined by the bindings in \c{envir}.

\begin{lstlisting}
> substitute(expression(a+b), list(a=1)))  
expression(1+b)
\end{lstlisting}

\noindent
R allows programmatic manipulation of parse trees, which are themselves
first class objects. They are evaluated using the \c{eval()}
function. There are three alternative forms  of \eval in the \c{base} package:
\c{evalq}, \c{eval.parent}, \c{local}.

\begin{lstlisting}
 eval(expr,
      envir = parent.frame(),
      enclos = if(is.list(envir) || is.pairlist(envir))
               parent.frame() else baseenv())
\end{lstlisting}

The arguments are as follows: \c{expr} is the value to be evaluated,
\c{envir} is the environment in which \c{expr} is to be evaluated, and
\c{enclos} is relevant when \c{envir} is a list or a data frame, then it
specifies where R looks for objects not found in \c{envir}. This can be
\c{NULL} (interpreted as the base package environment, \c{baseenv()}) or an
environment.

The semanitcs of \eval is to evaluate \c{expr} in the environment specified
by \c{envir} and return the computed value. If \c{envir} is not specified,
then the default is \c{parent.frame()} (the environment where the call to
\eval was made).  Values to be evaluated can be of types \c{call} (AST for
function calls) or \c{expression} (AST nodes for arbitrary expressions) or
\c{name} (when the symbol is looked up in the current scope and its binding
is evaluated), a promise or any of the basic types such as vectors,
functions and environments (which are returned unchanged).

The \c{evalq} form is equivalent to \c{eval(quote(expr), ...)}. eval
evaluates its first argument in the current scope before passing it to the
evaluator: evalq avoids this.

\c{eval.parent(expr, n)} is a shorthand for \c{eval(expr, parent.frame(n))}.

If envir is a list (such as a data frame) or pairlist, it is copied into a
temporary environment (with enclosure enclos), and the temporary environment
is used for evaluation. So if expr changes any of the components named in
the (pair)list, the changes are lost.

If envir is NULL it is interpreted as an empty list so no values could be
found in envir and look-up goes directly to enclos.

\c{local} evaluates an expression in a local environment. It is equivalent to
evalq except that its default argument creates a new, empty
environment. This is useful to create anonymous recursive functions and as a
kind of limited namespace feature since variables defined in the environment
are not visible from the outside.

@@

Unlike CRAN, Kaggle is not a curated repository and therefore there are no
guarantees about the quality of the code. After downloading all of the
\CorpusKaggle kernels and extracting the R code from the various
formats\footnote{We use \c{rmarkdown} package to convert from Jupyer
  notebook to Rmarkdown as well as to convert from Rmarkdown to plain R}, we
found that \CorpusDuplicatedKaggle were whole-file duplicates. From
\CorpusRunnableKaggle kernels, \CorpusFailedKaggle failed to execute. Next to
various runtime exceptions, common problems were parse errors and misspelled
package names.
